---
title: "Number of Foster Youth by Schools and Districts in Massachusetts"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
  html_notebook:
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---

<style type="text/css">

body{ /* Normal  */
      font-size: 14px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 18px;
}
h1 { /* Header 1 */
  font-size: 18px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 16px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>


In this report the number of students with foster care for all grades (03-08 and 10) are studied for school year 2019 in all schools and districts in Massachusetts state.\

# Foster Youth Populations Based on MCAS Math Scores by Schools

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
foster <- read.csv("~/Desktop/NextGenMCAS.schools.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)

##### MATH ######
math <- foster[foster$Subject == "MATH",]
math <- math[-24,]
math$No..of.Students.Included <- as.numeric(math$No..of.Students.Included)
schools <- aggregate(No..of.Students.Included ~ School.Name, data = math, sum)
schools
for (i in 1:nrow(schools)) {
  schools$School.Name[i] <-  gsub("^.*?- ","",schools$School.Name[i])}

schools$School.Name <- gsub("Mildred Avenue K-8", "Mildred Avenue K-8 School", schools$School.Name)
schools$School.Name <- gsub("William McKinley",  "William McKinley School", schools$School.Name)
schools$School.Name <- gsub("Dupont Middle",  "Dupont Middle School", schools$School.Name)
schools$School.Name <- gsub("Northwest",  "Northwest School", schools$School.Name)
schools$School.Name <- gsub("John T. Nichols Middle",  "John T Nichols Middle School", schools$School.Name)
schools$School.Name <- gsub("Northbridge Middle",  "Northbridge Middle School", schools$School.Name)
schools$School.Name <- gsub("Randolph Community Middle",  "Randolph Community Middle School", schools$School.Name)
schools$School.Name <- gsub("Forest Park Middle",  "Forest Park Middle School", schools$School.Name)
schools$School.Name <- gsub("John F Kennedy Middle",  "John F Kennedy Middle School", schools$School.Name)
schools$School.Name <- gsub("John J Duggan Middle",  "John J Duggan Middle School", schools$School.Name)
schools$School.Name <- gsub("M Marcus Kiley Middle",  "M. Marcus Kiley Middle School", schools$School.Name)
schools$School.Name <- gsub("Benjamin Friedman Middle",  "Benjamin Friedman Middle School", schools$School.Name)
schools$School.Name <- gsub("Wareham Middle",  "Wareham Middle School", schools$School.Name)
schools$School.Name <- gsub("Forest Grove Middle",  "Forest Grove Middle School", schools$School.Name)
schools$School.Name <- gsub("Knox Trail Middle School",  "Knox Trail Junior High School", schools$School.Name)
names(schools)[2] <- 'Number.FosterStudents'
```

\

## The Graph for Foster Youth Populations Based on MCAS Math Scores by Schools
\

Followed is the barplot for total number of students with foster care according to MCAS Math for school year 2019 by schools in Massachusetts.\
\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
library(ggplot2)
library(dplyr)
library(forcats)
schools %>%
  mutate(School.Name = fct_reorder(School.Name, desc(Number.FosterStudents))) %>%
ggplot(aes(x=School.Name, y=Number.FosterStudents)) +
  geom_bar(stat="identity", fill="navy")+
  geom_text(aes(label=Number.FosterStudents), vjust=-0.3, size=3.5) + theme_bw() + xlab("School Name") + ylab("Population of Foster Youth Based on MCAS Math") + ggtitle("Number of Foster Care Students by Schools in MA in 2019 for All Grades According to MCAS Math Scores") + theme(text = element_text(size=8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8)) 
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
library(sf)
library(tmap)
schools <- transform(schools, FosterStudents_Percent = schools$Number.FosterStudents/(sum(schools$Number.FosterStudents))*100)
schools
point <- st_read(dsn="~/Desktop/School_data_CFJJ_analysis_for_Mehrnaz/Shapefiles_SCHOOL_SCHOOLDISTRICT/SCHOOLS_PT.shp")
point <- merge(point, schools, by.x='NAME', by.y = 'School.Name', all.y=TRUE)
point <- point[-c(10,11,12),]
allStudents <- read.csv("~/Desktop/allStudents.District.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
poly <- st_read(dsn="~/Desktop/School_data_CFJJ_analysis_for_Mehrnaz/Shapefiles_SCHOOL_SCHOOLDISTRICT/SCHOOLDISTRICTS_POLY.shp")
poly <- merge(poly, allStudents, by.x='DISTRICT_N', by.y = 'District.Name', all = TRUE)
```

\

## The Map for Foster Youth Populations Based on MCAS Math Scores by Schools
\

Followed is the map, which indicates the number of students with foster care based on MCAS Math by schools in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
tmap_mode('view')
tm_shape(poly) + tm_fill('grey86') + tm_borders("gold", lwd = 1) + tm_basemap(server = c('OpenStreetMap')) +
  tm_shape(point) + tm_dots("Number.FosterStudents", style = "jenks", size = 0.3, palette='Greens')  +  tm_view(alpha = 0.85, view.legend.position = c("left", "bottom")) + tm_layout(title = 'Population of Foster Youth Based on MCAS Math by Schools')
```


\

## The Map for Percentage of Foster Youth Populations Based on MCAS Math Scores by Schools
\

Followed is the map, which indicates the percentage of students with foster care based on MCAS Math by schools in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
tmap_mode('view')
tm_shape(poly) + tm_fill("antiquewhite") + tm_borders("antiquewhite4", lwd = 1) + tm_basemap(server = c('OpenStreetMap')) +
  tm_shape(point) + tm_dots("FosterStudents_Percent", style = "equal", size = 0.3, palette='Blues', popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  +  tm_view(alpha = 0.85, view.legend.position = c("left", "bottom")) 
```



\

# Foster Youth Populations Based on MCAS ELA Scores by Schools
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
ELA <- foster[foster$Subject == "ELA",]
ELA <- ELA[-27,]
ELA$No..of.Students.Included <- as.numeric(ELA$No..of.Students.Included)
schools2 <- aggregate(No..of.Students.Included ~ School.Name, data = ELA, sum)

for (i in 1:nrow(schools2)) {
  schools2$School.Name[i] <-  gsub("^.*?- ","",schools2$School.Name[i])}

schools2$School.Name <- gsub("Mildred Avenue K-8", "Mildred Avenue K-8 School", schools2$School.Name)
schools2$School.Name <- gsub("William McKinley",  "William McKinley School", schools2$School.Name)
schools2$School.Name <- gsub("Dupont Middle",  "Dupont Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Northwest",  "Northwest School", schools2$School.Name)
schools2$School.Name <- gsub("John T. Nichols Middle",  "John T Nichols Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Northbridge Middle",  "Northbridge Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Randolph Community Middle",  "Randolph Community Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Forest Park Middle",  "Forest Park Middle School", schools2$School.Name)
schools2$School.Name <- gsub("John F Kennedy Middle",  "John F Kennedy Middle School", schools2$School.Name)
schools2$School.Name <- gsub("John J Duggan Middle",  "John J Duggan Middle School", schools2$School.Name)
schools2$School.Name <- gsub("M Marcus Kiley Middle",  "M. Marcus Kiley Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Benjamin Friedman Middle",  "Benjamin Friedman Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Wareham Middle",  "Wareham Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Forest Grove Middle",  "Forest Grove Middle School", schools2$School.Name)
schools2$School.Name <- gsub("Knox Trail Middle School",  "Knox Trail Junior High School", schools2$School.Name)
names(schools2)[2] <- 'Number.FosterStudents'
```

\

## The Graph for Foster Youth Populations Based on MCAS ELA Scores by Schools
\

Followed is the barplot for total number of students with foster care according to MCAS ELA for school year 2019 by schools in Massachusetts.\
\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
schools2 %>%
  mutate(School.Name = fct_reorder(School.Name, desc(Number.FosterStudents))) %>%
ggplot(aes(x=School.Name, y=Number.FosterStudents)) +
  geom_bar(stat="identity", fill="cadetblue")+
  geom_text(aes(label=Number.FosterStudents), vjust=-0.3, size=3.5) + theme_bw() + xlab("School Names") + ylab("Population of Foster Youth Based on MCAS ELA") + ggtitle("Number of Foster Care Students by Schools in MA in 2019 for All Grades According to MCAS ELA Scores") + theme(text = element_text(size=8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 10)) 
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
schools2 <- transform(schools2, FosterStudents_Percent = schools2$Number.FosterStudents/(sum(schools2$Number.FosterStudents))*100)
point <- st_read(dsn="~/Desktop/School_data_CFJJ_analysis_for_Mehrnaz/Shapefiles_SCHOOL_SCHOOLDISTRICT/SCHOOLS_PT.shp")
point <- merge(point, schools2, by.x='NAME', by.y = 'School.Name', all.y=TRUE)
point <- point[-c(9,11,12),]
```

\

## The Map for Foster Youth Populations Based on MCAS ELA Scores by Schools
\

Followed is the map, which indicates the number of students with foster care based on MCAS ELA by schools in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
tmap_mode('view')
tm_shape(poly) + tm_fill('grey86') + tm_borders("gold", lwd = 1) + tm_basemap(server = c('OpenStreetMap')) +
  tm_shape(point) + tm_dots("Number.FosterStudents", style = "jenks", size = 0.3)  +  tm_view(alpha = 0.85, view.legend.position = c("left", "bottom")) 
```


\

## The Map for Percentage of Foster Youth Populations Based on MCAS ELA Scores by Schools
\

Followed is the map, which indicates the percentage of students with foster care based on MCAS ELA by schools in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
tmap_mode('view')
tm_shape(poly) + tm_fill("antiquewhite") + tm_borders("antiquewhite4", lwd = 1) + tm_basemap(server = c('OpenStreetMap')) +
  tm_shape(point) + tm_dots("FosterStudents_Percent", style = "equal", size = 0.3,  palette='Purples', popup.format = list(fun = make_percent), legend.format = list(fun = make_percent))  +  tm_view(alpha = 0.85, view.legend.position = c("left", "bottom")) 
```

\

# Foster Youth Populations Based on MCAS Math Scores by District
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
foster.Dist <- read.csv("~/Desktop/NextGenMCAS.District.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
foster.Dist10 <- read.csv("~/Desktop/NextGenMCAS.District.Grade10.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
foster.Dist <- rbind(foster.Dist, foster.Dist10)
foster.Dist <- foster.Dist[-c(158, 157, 144, 145, 146),]
foster.Dist$No..of.Students.Included <- as.numeric(foster.Dist$No..of.Students.Included)
dist.math <- foster.Dist[foster.Dist$Subject == "MATH",]
dist.math <- aggregate(No..of.Students.Included ~ District.Name, data = dist.math, sum)
names(dist.math)[2] <- 'Number.FosterStudents'
dist.math
```

\

## The Graph for Foster Youth Populations Based on MCAS Math by District
\

Followed is the barplot for total number of students with foster care for all grades (03-08 and 10) according to MCAS Math scores for school year 2019 by district in Massachusetts.\
\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
dist.math %>%
  mutate(District.Name = fct_reorder(District.Name, desc(Number.FosterStudents))) %>%
ggplot(aes(x=District.Name, y=Number.FosterStudents)) +
  geom_bar(stat="identity", fill="burlywood3")+ theme_bw() + xlab("District Name") + ylab("Population of Foster Youth Based on MCAS Math") + ggtitle("Number of Foster Care Students by Districts in MA in 2019 for All Grades According to MCAS Math Scores") + theme(text = element_text(size=8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8)) 
```

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
dist.math <- transform(dist.math, FosterStudents_Percent = dist.math$Number.FosterStudents/(sum(dist.math$Number.FosterStudents))*100)
polygon <- st_read(dsn="~/Desktop/School_data_CFJJ_analysis_for_Mehrnaz/Shapefiles_SCHOOL_SCHOOLDISTRICT/SCHOOLDISTRICTS_POLY.shp")
polygon <- merge(polygon, dist.math, by.x = 'DISTRICT_N', by.y='District.Name', all = TRUE)
```

\

## The Map for Foster Youth Populations Based on MCAS Math Scores by District
\

Followed is the map, which indicates the number of students with foster care for all grades (03-08 and 10) based on MCAS Math scores by districts in MA for school year 2019.\


```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
tmap_mode('view')
tm_shape(polygon) + tm_polygons('Number.FosterStudents', style = "jenks",  palette='Reds')  + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_basemap(server = c('OpenStreetMap')) 
```


\

## The Map for Percentage of Foster Youth Populations Based on MCAS Math Scores by Districts
\

Followed is the map, which indicates the percentage of students with foster care for all grades (03-08 and 10) based on MCAS Math scores by districts in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
tmap_mode('view')
tm_shape(polygon) + tm_polygons('FosterStudents_Percent', style = "jenks", palette='Blues', popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_basemap(server = c('OpenStreetMap')) 
```




\

# Foster Youth Populations Based on MCAS ELA Scores by District
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
foster.Dist <- read.csv("~/Desktop/NextGenMCAS.District.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
foster.Dist10 <- read.csv("~/Desktop/NextGenMCAS.District.Grade10.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
foster.Dist <- rbind(foster.Dist, foster.Dist10)
foster.Dist <- foster.Dist[-c(158, 157, 144, 145, 146),]
foster.Dist$No..of.Students.Included <- as.numeric(foster.Dist$No..of.Students.Included)
dist.ELA <- foster.Dist[foster.Dist$Subject == "ELA",]
dist.ELA <- aggregate(No..of.Students.Included ~ District.Name, data = dist.ELA, sum)
names(dist.ELA)[2] <- 'Number.FosterStudents'
dist.ELA
```

\

## The Graph for Foster Youth Populations Based on MCAS ELA Scores by District
\

Followed is the barplot for total number of students with foster care for all grades (03-08 and 10) according to MCAS ELA scores for school year 2019 by district in Massachusetts.\
\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
dist.ELA %>%
  mutate(District.Name = fct_reorder(District.Name, desc(Number.FosterStudents))) %>%
ggplot(aes(x=District.Name, y=Number.FosterStudents)) +
  geom_bar(stat="identity", fill="firebrick4")+ theme_bw() + xlab("District Name") + ylab("Population of Foster Youth Based on MCAS ELA") + ggtitle("Number of Foster Care Students by Districts in MA in 2019 for All Grades According to MCAS ELA Scores") + theme(text = element_text(size=8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8)) 
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
dist.ELA <- transform(dist.ELA, FosterStudents_Percent = dist.ELA$Number.FosterStudents/(sum(dist.ELA$Number.FosterStudents))*100)
polygon <- st_read(dsn="~/Desktop/School_data_CFJJ_analysis_for_Mehrnaz/Shapefiles_SCHOOL_SCHOOLDISTRICT/SCHOOLDISTRICTS_POLY.shp")
polygon <- merge(polygon, dist.ELA, by.x = 'DISTRICT_N', by.y='District.Name', all = TRUE)
```

\

## The Map for Foster Youth Populations Based on MCAS ELA Scores by District
\

Followed is the map, which indicates the number of students with foster care for all grades (03-08 and 10) based on MCAS ELA scores by districts in MA for school year 2019.\


```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
tmap_mode('view')
tm_shape(polygon) + tm_polygons('Number.FosterStudents', style = "jenks",  palette='Greens')  + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_basemap(server = c('OpenStreetMap')) 
```


\

## The Map for Percentage of Foster Youth Populations Based on MCAS ELA Scores by Districts
\

Followed is the map, which indicates the percentage of students with foster care for all grades (03-08 and 10) based on MCAS ELA scores by districts in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
tmap_mode('view')
tm_shape(polygon) + tm_polygons('FosterStudents_Percent', style = "jenks", popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_basemap(server = c('OpenStreetMap')) 
```




\

# Foster Youth Populations Based on MCAS SCI Scores by District
\

```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
foster.Dist <- read.csv("~/Desktop/NextGenMCAS.District.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
foster.Dist10 <- read.csv("~/Desktop/NextGenMCAS.District.Grade10.csv", sep = ",", header = TRUE, stringsAsFactors=FALSE)
foster.Dist <- rbind(foster.Dist, foster.Dist10)
foster.Dist <- foster.Dist[-c(158, 157, 144, 145, 146),]
foster.Dist$No..of.Students.Included <- as.numeric(foster.Dist$No..of.Students.Included)
dist.SCI <- foster.Dist[foster.Dist$Subject == "SCI",]
dist.SCI <- aggregate(No..of.Students.Included ~ District.Name, data = dist.SCI, sum)
names(dist.SCI)[2] <- 'Number.FosterStudents'
```

\

## The Graph for Foster Youth Populations Based on MCAS SCI Scores by District
\

Followed is the barplot for total number of students with foster care for all grades according to MCAS SCI scores for school year 2019 by district in Massachusetts.\
\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
dist.SCI %>%
  mutate(District.Name = fct_reorder(District.Name, desc(Number.FosterStudents))) %>%
ggplot(aes(x=District.Name, y=Number.FosterStudents)) +
  geom_bar(stat="identity", fill="deeppink4")+ theme_bw() + xlab("District Name") + ylab("Population of Foster Youth Based on MCAS SCI") + ggtitle("Number of Foster Care Students by District in MA in 2019 for All Grades According to MCAS SCI Scores") + theme(text = element_text(size=8)) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8)) + geom_text(aes(label=Number.FosterStudents), vjust=-0.3, size=3.5) 
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results='hide', comment=""}
dist.SCI <- transform(dist.SCI, FosterStudents_Percent = dist.SCI$Number.FosterStudents/(sum(dist.SCI$Number.FosterStudents))*100)
polygon <- st_read(dsn="~/Desktop/School_data_CFJJ_analysis_for_Mehrnaz/Shapefiles_SCHOOL_SCHOOLDISTRICT/SCHOOLDISTRICTS_POLY.shp")
polygon <- merge(polygon, dist.SCI, by.x = 'DISTRICT_N', by.y='District.Name', all = TRUE)
```

\

## The Map for Foster Youth Populations Based on MCAS SCI Scores by District
\

Followed is the map, which indicates the number of students with foster care for all grades based on MCAS SCI scores by districts in MA for school year 2019.\


```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
tmap_mode('view')
tm_shape(polygon) + tm_polygons('Number.FosterStudents', style = "jenks",  palette='Oranges')  + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_basemap(server = c('OpenStreetMap')) 
```


\

## The Map for Percentage of Foster Youth Populations Based on MCAS SCI Scores by Districts
\

Followed is the map, which indicates the percentage of students with foster care for all grades based on MCAS SCI scores by districts in MA for school year 2019.\

```{r echo=FALSE, warning=FALSE, message=FALSE, comment=""}
make_percent <- function(x, digits = 2) {
  paste0("%", formatC(x, digits = digits, format = "f", big.mark = ","))
}
tmap_mode('view')
tm_shape(polygon) + tm_polygons('FosterStudents_Percent', style = "jenks", palette='Purples', popup.format = list(fun = make_percent), legend.format = list(fun = make_percent)) + tm_view(alpha = 0.85,view.legend.position = c("left", "bottom"),basemaps = "Stamen.TonerLite") + tm_borders(col = "darkgray") + tmap_options(show.messages = FALSE) + tm_basemap(server = c('OpenStreetMap')) 
```